### 2016-12-30 Chapters/experiments.tex
First full draft of chap3 suppinfo

We applied this experimental set-up to the reconstruction of the resources allocation strategy employed by \textit{E. coli} during a nutrient upshift, with the aim to observe the bang-bang profile predicted in Chapter~\ref{chap:theory}.
We developed a Kalman smoothing procedure adapted to our question and showed how it can be implemented to reconstruct the internal state of a cell from time-lapse microscopy images.
Similar to hidden Markov chains, Kalman filtering and smoothing are powerful techniques for the reconstruction of unobserved signals from noisy measurements, and have had numerous applications in a variety of domains, though less in biology than in other fields.
We showed in this chapter that the Bayesian framework of Kalman smoothing can be exploited for growth rate reconstruction from a probabilistic prior defining the relation between mother and daughter cells.
We also developed another variant of Kalman smoothing for the reconstruction of the resource allocation variable $\alpha (\cdot)$ that is at the heart of the models of Chapter~\ref{chap:theory}.
Our work on synthetic data showed that, despite the challenge posed by the reconstruction of a discontinuous signal, the algorithm was capable of recovering an oscillatory pattern close to the expected on-off profile.
Unfortunately, its application to the real data did not allow an unambiguous conclusion to be drawn: while oscillations do occur, and a first peak between 0 and 120 min after the upshift is visible in almost all of the cells, it turned out to be far from straightforward to distinguish a real oscillatory pattern from artifacts of the data analysis and signal reconstruction methods.
Below we propose several improvements that will be addressed in future work.

@@

It is also possible to optimize the image analysis techniques used in this study.
We segmented the images using the most powerful and ubiquitous segmentation algorithm available: the human brain.
In other words, we manually selected the pixels representing the poles of a single bacteria per well on the fluorescence images, and used this information to arbitrarily define a rectangle around the cell of interest.
While this is far from satisfying, it may turn out to be difficult to improve upon this, as it seems that the first rule of image analysis is that every application is more or less unique.
Every algorithm has to make assumptions about the object it is trying to recognize.
For imaging of microbial cells, it could be the curvature at the poles, the size of the pixels, the homogeneity of the background, the homogeneity of intracellular fluorescence, ...
While some parameters can be tweaked, some assumptions are always hard-coded in the algorithm and reflect the philosophy used to address the identification problem.
Since the preliminary analysis reported here was completed, we started a collaboration with the authors of FluoBacTracker~\cite{fluobactracker} that focuses on adapting their software to our set-up in a near future.
It should provide a robust, automated way to identify and track the cells, enhancing the reproducibility of the analysis and increasing the number of cells available for computing population statistics.

@@

\begin{figure}
\includegraphics[width=13cm]{/home/geiselmann/Persons/Nils/Thesis/Man/Fig/NilsCloning}\caption{\label{fig:Cloning}Construction of the gfp-tagged ribosome subunit.
The translational fusion of \textit{gfp\_mut2} with \textit{rpsB}
was constructed on the chromosome of \textit{E.~coli} (top line)
by homologous recombination of \textit{gfp\_mut2} followed by a selection
``cassette'' (second line). A small linker was inserted between
\textit{rpsB} and \textit{gfp\_mut2} in order to minimize interference
of the fluorescent protein with ribosome functioning. The selection
cassette consists of a positive selection marker, the gene coding
for the resistance to the antibiotic kanamycine, and a negative selection
marker, the gene coding for the toxin CcdB. The latter is transcribed
from the $p_{BAD}$ promoter, which is only activated in the presence
of arabinose in the culture medium. Homologous recombination is indicated
by the grey shaded lines. The resulting recombination product (third
line) contains the desired fusion protein followed by the selection
cassette on the chromosome of the bacterium. A second homologous recombination,
using an oligonucleotide complementary to the end of \textit{gfp\_mut2}
and the beginning the original region downstream of \textit{rpsB}
removes the selection cassette. The resulting strain (line four) carries
the translational fusion of \textit{gfp\_mut2} to \textit{rpsB} without
any other modification of the chromosome.}

\selectlanguage{english}%
\selectlanguage{french}%
\end{figure}

The DNA fragment containing the selection cassette was amplified using
two long primers annealing, respectively, to the region just before
the STOP codon of \textit{rpsB}, and to the beginning of \textit{gfpmut2}.
The selection cassette contains an antiobiotic resistance gene, kanamycin
(positive selection), and a gene coding for the CcdB toxin under the
control of the P\textsubscript{BAD} promoter (negative selection
in presence of arabinose). This cassette is referred below as \textit{kan-p\textsubscript{BAD}-ccdB}.

Another DNA fragment containing \textit{gfp\_mut2}~\cite{zaslaver_comprehensive_2006}
without the ATG start codon was amplified using long primers annealing,
respectively, to the C-terminus of \ref{chap:theory}\textit{rpsB}
(just before the STOP codon), and the end of the \textit{kan-P\textsubscript{BAD}-ccdB}
cassette. The first primer also contained a 18-bp (base pair) linker,
inserting the same six amino-acids that Bakshi had used for their
construction ~\cite{bakshi_superresolution_2012}.

The two fragments, the \textit{gfp\_mut2} reporter including the linker
and the \textit{kan-P\textsubscript{BAD}-ccdB} cassette, were assembled
using Gibson assembly~\cite{gibson_enzymatic_2009} using a commercial
Gibson Assembly mix (New England Biolabs). A final product of 2683
bp was obtained: 
\begin{itemize}
\item {*}(50 bp) the C-terminus of \textit{rpsB} without the STOP codon 
\item (18 bp) a linker (see Supporting Information) 
\item (714 bp) the \textit{gfpmut2} sequence without the initial ATG 
\item (1851 bp) the \textit{kan-P\textsubscript{BAD}-ccdB} cassette 
\item {*}(50 bp) the region directly after \textit{rpsB} in \textit{E.~coli} 
\end{itemize}
Regions labeled with {*} anneal to the \textit{E.~coli} chromosome.
The complete sequence of this fragment, as well as the sequence of
all the primers,are listed in the Supporting Information of this chapter.

This fragment was electroporated into the wildtype strain, BW25113,
containing the pSIM5 plasmid for lambda-red recombinaison \cite{sharan2009recombineering}.
A kanamycin-resistant colony was selected on LA-glucose medium and
verified to show green fluorescence in a Tecan microplate reader.
A 100-bp oligonucleotide, containing 50 bp of the end of \textit{gfpmut2}
and 50 bp of the region just downstream of \textit{rpsB} was electroporated
into this strain (sequence available in Supporting Information). The
homologous recombination removes the \textit{kan-P\textsubscript{BAD}-ccdB}
cassette. A colony was selected on an LA-arabinose plate. The clone
was tested for kanamycin-sensitivity and  green fluorescence. Finally,
the strain was grown overnight at 42$^{\circ}$C to get rid of the
pSIM5 plasmid, containing a temperature-sensitive origin of replication.
A chloramphenicol-sensitive colony was chosen andthe region after
\textit{rpsB} was verified by sequencing (full sequence available
in Supporting Information).

In parallel, the same protocol was used to construct \textit{mCherry}
and \textit{cfp} variants of the same strain. However, only the \textit{gfp\_mut2}
and \textit{mCherry} versions were successfully obtained. The full
sequence of the final \textit{rpsB-mCherry} strain is available in
Supporting Information.

The \textit{rpsB-gfp\_mut2} and \textit{rpsB-mCherry} strains were
characterized on different media using a Tecan microplate reader.
They possess a wild-type growth rate and sufficient fluorescence to
allow quantification. However, the \textit{rpsB-mCherry} strain exhibited
strange fluorescence dynamics, especially during growth transitions,
which made it unsuitable for our study. We therefore concentrated
our efforts on the \textit{rpsB-gfp} strain. Detaied data about the
characterization of the strain are available in Supporting Information.

@@

All experiments were carried out in sterilized media. Sterilization
was performed by autoclaving for instruments and filtration for solutions
(0.2 $\mu$m). During cloning, bacteria were grown in 20-mL Erlenmeyer
flasks filled with LB or spread on Petri dishes with LA. In all the
experiments, we used M9 minimal medium supplemented with trace elements,
thiamine and a carbon source. The full recipe is reproduced below.
Numbers in squared brackets are characteristics of the stock solution.
All stock solutions were stored at room temperature, except for FeSO$_{4}$~{[}30
g.L\textsuperscript{-1}{]} and Thiamine~{[}10 g.L\textsuperscript{-1}{]},
which were stored respectively at -20$^{\circ}$C and 4$^{\circ}$C. 
\begin{center}
\begin{tabular}{|l r|}
\hline 
\multicolumn{2}{|c}{\textbf{M9 medium (100 mL)}} & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
CaCl2 {[}1 mol.L\textsuperscript{-1}{]}  & 10 $\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
MgSO$_{4}$ {[}1 mol.L\textsuperscript{-1}{]}  & 200 $\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
5x Salts  & 20 mL  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
Traces elements  & 90 $\mu$l  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
FeSO$_{4}$ {[}30 g.L\textsuperscript{-1}{]}  & 10 $\mu$l  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
Thiamine {[}10 g.L\textsuperscript{-1}{]}  & 50 $\mu$l  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
Carbon source  & at will  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
H$_{2}$0 {[}18.2 M$\Omega$.cm{]}  & to 100 mL  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
\end{tabular}%
\begin{tabular}{|l r|}
\hline 
\multicolumn{2}{|c}{\textbf{Traces elements (0.9 mL)}} & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
H$_{2}$0 {[}18.2 M$\Omega$.cm{]}  & 200$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
Na$_{2}$EDTA 2H$_{2}$O {[}150 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
ZnSO$_{4}$ 7H$_{2}$O {[}45 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
CoCl$_{2}$ 6H$_{2}$O {[}3 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
MnCl$_{2}$ 4H$_{2}$O {[}10 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
H$_{3}$BO$_{3}$ {[}10 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
Na$_{2}$MoO$_{4}$ 2H$_{2}$O {[}4 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
CuSO$_{4}$ 5H$_{2}$O {[}3 g.L\textsuperscript{-1}{]}  & 100$\mu$l & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
\end{tabular}%
\begin{tabular}{|l r|}
\hline 
\multicolumn{2}{|c}{\textbf{5x Salts (100 mL)}} & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
Na$_{2}$HPO$_{4}$2H$_{2}$0  & 4.25 g  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
KH$_{2}$HPO$_{4}$  & 1.5 mg  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
NaCl  & 0.25 g & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
NH$_{4}$Cl & 0.5 g & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
H$_{2}$0 {[}18.2 M$\Omega$.cm{]}  & to 100 mL  & \selectlanguage{english}%
\selectlanguage{french}%
\tabularnewline
\hline 
\end{tabular}
\par\end{center}

A large quantity of M9 was prepared several days before the experiment,
and stored at 4$^{\circ}$C. It contained all the necessary components
except FeSO$_{4}$, thiamine, and carbon sources, which were added
just before inocculation of the pre-culture. Except where stated otherwise,
the growth temperature was 37$^{\circ}$C. M9 Acetate contains 0.2\%
Acetate (in mass of C$_{2}$H$_{3}$O$_{2}$ per mass of solution),
and M9 Glucose contains 0.2\% Glucose (in mass of D-(+)-Glucose per
mass of solution).

Four days before the measurements, the glycerol stock containing the
\textit{rpsB-gfp} strain was spread on a Petri dish of LA, and incubated
at 37$^{\circ}$C. On the following day, an isolated colony was was
used to inoculate a cotton-plugged, sterile flask containing 20 mL
of M9 acetate (Fig.~\ref{fig:experiment_schema}). The time of inoculation
was calculated such that the culture obtains an OD of 0.3-0.4 OD at
the beginning of the experiment. This optical density corresponds
to a culture in mid-exponential growth phase.

At the day of the experiment, the preculture was concentrated by centrifugation
and re-suspended in 5 mL of M9 acetate, supplemented with 50 mg.mL\textsuperscript{-1}
BSA (passivation buffer) for injection into the microfluidic device.
Channels were filled by diffusion until most of them contained cells
($\sim$1 hour). Data acquisition started after a constant medium
flow was successfully obtained ($\sim$1 hour, depending on the quality
of the microfluidics device).

@@

The device was connected using 0.023\textquotedbl{} inner diameter
polyethylene tubes to a waste and a sterile bottle containing 200
mL of growth medium, which was enough for several days of acquisition.
A microfluidics pump (Elvesys) containing an output flow sensor module
was plugged to the medium bottle and a pressure up to 2~bars was
applied to the bottle. The output flow was set to 50 $\mu$L.min$^{-1}$.

@@

In order to register consecutive image of a time series, we calculated
the offset between consecutive images of a given field using cross-correlation
in the Fourier space~\cite{guizar-sicairos_efficient_2008} (see
Scikit-Image documentation at \cite{skimage_cross-correlation}).
All images for a field were thereby aligned to the first acquisition
by a simple translation.

@@

The autofluorescence of the bacteria and the background fluorescence
of the medium were supposed to be negligible and were thus not corrected.
As a control, the same device was used to image bacteria in stationary
phase that do not produce any fluorescent protein. They were indistinguishable
from the camera noise. Of course, using this control we can not affirm
that the autofluorescence is also negligible when the bacteria are
actively growing. Furthermore, the autofluorescencecould change in
different media of interest, and may even be different in steady state
and during growth transitions. An independent estimate of the autofluorescence
is therefore difficult to obtain. In the section~\ref{sec:chap3_discussion},
we discuss possible improvements of the experimental setup that would
allow to dynamically co-estimate the autofluorescence with the ribosome
abundance in a single experiment.

@@

The data of the length and the RFU/pixel of each bacteria were analyzed
using Python~3.5.2. In particular, we manipulated the data using
Pandas~0.19.1, pykalman~0.9.5 and the curves submodule of wellfare~0.1.1.

Historical details about the Kalman smoothing procedure are reported
in section~\ref{sec:res_kalman} of the main text. We used the Additive
Unscented Kalman Filter implementation of the pykalman python module.
This class is reported to be more stable and computationally efficient
with non-linear problems containing additive noise.

As described in section~\ref{sec:res_kalman}, the reconstruction
of the growth rate was done on continuous portions of the length,
\textit{i.e.} between cell divisions. The full problem, as described
in Eq.~\ref{eq:full_mu_prob}, is reproduced below for clarity: 
\begin{eqnarray*}
\dot{V_{\lambda}}(t) & = & \mu(t)\cdot V_{\lambda}(t),\\
\dot{\mu}(t) & = & v(t),\\
\dot{v}(t) & = & w(t),
\end{eqnarray*}
with the measurement model: 
\[
L(t_{k})=V_{\lambda}(t_{k})+\epsilon_{k}.
\]
The parameters used as priors for the reconstruction of $\mu$ are
described below. We used an observation variance of 9 pixels\textsuperscript{2}
for the length $L$. The transition variance $\theta^{2}$ (\textit{a.k.a.}
the smoothing factor for $\mu$) is fixed at $10^{-8}$~min\textsuperscript{-6}
for the entire time series. Inheritance between mother and daughter
cells is taken into account by systematically choosing an initial
mean equal to the last estimated value for $\mu$, and to half the
last estimated value for $V_{\gamma}$ (as expected for a completely
symmetrical division). At the beginning of the experiment, when no
mother cells are available, these values were fixed to 15~pixels
for $V_{\gamma}$ and 0.004~min\textsuperscript{-1} for $\mu$.
The variances associated with these means are 16~pixels\textsuperscript{2}
and $10^{-4}$~min\textsuperscript{-2}, respectively, for $V_{\gamma}$
and $\mu$. The initial mean of $v$ is always taken as null, with
an initial variance of $10^{-8}$~min\textsuperscript{-6}. All the
cross-covariances are set to 0 because the system variables are independent
by construction.

>>>>

We applied this experimental setup to the reconstruction of the resources allocation strategy employed by \textit{E. coli} during a nutrient upshift, with the aim to observe the bang-bang profile predicted in Chapter~\ref{chap:theory}.
We developed a Kalman smoothing procedure adapted to our question and showed how it can be implemented to reconstruct the internal state of a cell from time-lapse microscopy images.
Similar to hidden Markov chains, Kalman filtering and smoothing are powerful techniques for the reconstruction of unobserved signals from noisy measurements, and have had numerous applications in a variety of domains, though less in biology than in other fields.
We showed in this chapter that the Bayesian framework of Kalman smoothing can be exploited for growth rate reconstruction from a probabilistic prior defining the relation between mother and daughter cells.
We also developed another variant of Kalman smoothing for the reconstruction of the resource allocation variable $\alpha (\cdot)$ that is at the heart of the models of Chapter~\ref{chap:theory}.
Our work on synthetic data showed that, despite the challenge posed by the reconstruction of a discontinuous signal, the algorithm was capable of recovering an oscillatory pattern close to the expected on-off profile.
Unfortunately, its application to the real data did not allow an unambiguous conclusion to be drawn: while oscillations do occur, and a first peak between 0 and 120 min after the upshift is visible in almost all of the cells, it turned out to be far from straightforward to distinguish a real oscillatory pattern from artifacts of the data analysis and signal reconstruction methods.
Below we propose several improvements that will be addressed in future work.

@@

It is also possible to optimize the image analysis techniques used in this study.
We segmented the images using the most powerful and ubiquitous segmentation algorithm available: the human brain.
In other words, we manually selected the pixels representing the poles of a single bacteria per well on the fluorescence images, and used this information to arbitrarily define a rectangle around the cell of interest.
While this is far from satisfying, it may turn out to be difficult to improve upon this, as it seems that the first rule of image analysis is that every application is more or less unique.
Every algorithm has to make assumptions about the object it is trying to recognize.
For imaging of microbial cells, it could be the curvature at the poles, the size of the pixels, the homogeneity of the background, the homogeneity of intracellular fluorescence, ...
While some parameters can be tweaked, some assumptions are always hard-coded in the algorithm and reflect the philosophy used to address the identification problem.
Since the preliminary analysis reported here was completed, we started a collaboration with the authors of FluoBacTracker~\cite{fluobactracker} that focuses on adapting their software to our setup in a near future.
It should provide a robust, automated way to identify and track the cells, enhancing the reproducibility of the analysis and increasing the number of cells available for computing population statistics.

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/meth_cloning}
\caption{
\textbf{Construction of the gfp-tagged ribosome subunit.}
The translational fusion of \textit{gfp\_mut2} with \textit{rpsB}
was constructed on the chromosome of \textit{E.~coli} (top line)
by homologous recombination of \textit{gfp\_mut2} followed by a selection
``cassette'' (second line). A small linker was inserted between
\textit{rpsB} and \textit{gfp\_mut2} in order to minimize interference
of the fluorescent protein with ribosome functioning. The selection
cassette consists of a positive selection marker, the gene coding
for the resistance to the antibiotic kanamycine, and a negative selection
marker, the gene coding for the toxin CcdB. The latter is transcribed
from the $p_{BAD}$ promoter, which is only activated in the presence
of arabinose in the culture medium. Homologous recombination is indicated
by the grey shaded lines. The resulting recombination product (third
line) contains the desired fusion protein followed by the selection
cassette on the chromosome of the bacterium. A second homologous recombination,
using an oligonucleotide complementary to the end of \textit{gfp\_mut2}
and the beginning the original region downstream of \textit{rpsB}
removes the selection cassette. The resulting strain (line four) carries
the translational fusion of \textit{gfp\_mut2} to \textit{rpsB} without
any other modification of the chromosome.}
\label{fig:Cloning}
\end{figure}

The DNA fragment containing the selection cassette was amplified using
two long primers annealing, respectively, to the region just before
the STOP codon of \textit{rpsB}, and to the beginning of \textit{gfp\_mut2}.
The selection cassette contains an antiobiotic resistance gene, kanamycin
(positive selection), and a gene coding for the CcdB toxin under the
control of the P\textsubscript{BAD} promoter (negative selection
in presence of arabinose). This cassette is referred below as \textit{kan-p\textsubscript{BAD}-ccdB}.

Another DNA fragment containing \textit{gfp\_mut2}~\cite{zaslaver_comprehensive_2006}
without the ATG start codon was amplified using long primers annealing,
respectively, to the C-terminus of \textit{rpsB}
(just before the STOP codon), and the end of the \textit{kan-p\textsubscript{BAD}-ccdB}
cassette. The first primer also contained a 18-bp (base pair) linker,
inserting the same six amino-acids that Bakshi \textit{et al.} have used for their
construction ~\cite{bakshi_superresolution_2012}.

The two fragments, the \textit{gfp\_mut2} reporter including the linker
and the \textit{kan-p\textsubscript{BAD}-ccdB} cassette, were assembled
by Gibson assembly~\cite{gibson_enzymatic_2009} using a commercial
Gibson Assembly mix (New England Biolabs). A final product of 2683
bp was obtained: 
\begin{itemize}
\item {*}(50 bp) the C-terminus of \textit{rpsB} without the STOP codon 
\item (18 bp) a linker
\item (714 bp) the \textit{gfpmut2} sequence without the initial ATG 
\item (1851 bp) the \textit{kan-p\textsubscript{BAD}-ccdB} cassette 
\item {*}(50 bp) the region directly after \textit{rpsB} in \textit{E.~coli} 
\end{itemize}
Regions labeled with {*} anneal to the \textit{E.~coli} chromosome.
The complete sequence of this fragment, as well as the sequences of
all the primers, are listed in \ref{S5_Text}.

This fragment was electroporated into the wildtype strain, BW25113,
containing the pSIM5 plasmid for lambda-red recombinaison~\cite{sharan_recombineering_2009}.
A kanamycin-resistant colony was selected on LA-glucose medium and
verified to show green fluorescence (485~nm excitation, 535~nm emission) in a microplate reader (Tecan Infinite 200 pro).
A 100-bp oligonucleotide, containing 50 bp of the end of \textit{gfp\_mut2}
and 50 bp of the region just downstream of \textit{rpsB} was electroporated
into this strain (sequence available in \ref{S5_Text}). The
homologous recombination removes the \textit{kan-p\textsubscript{BAD}-ccdB}
cassette. A colony was selected on an LA-arabinose plate. The clone
was tested for kanamycin-sensitivity and  green fluorescence. Finally,
the strain was grown overnight at 42$^{\circ}$C to get rid of the
pSIM5 plasmid, containing a temperature-sensitive origin of replication.
A chloramphenicol-sensitive colony was chosen and the region after
\textit{rpsB} was verified by sequencing (full sequence obtained available in \ref{S5_Text}).

In parallel, the same protocol was used to construct \textit{mCherry}
and \textit{cfp} variants of the same strain. However, only the \textit{gfp\_mut2}
and \textit{mCherry} versions were successfully obtained. The full
sequence of the final \textit{rpsB-mCherry} strain is available in \ref{S5_Text}.

The \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains were
characterized on different media using a Tecan microplate reader (\ref{S6_Text}, in particular Fig.~\ref{fig:suppinfo_caract} for growth on glucose).
They possess a wild-type growth rate and sufficient fluorescence to
allow quantification. However, the \textit{rpsB-mCherry} strain exhibited
strange fluorescence dynamics, especially during growth transitions,
which made it unsuitable for our study. We therefore concentrated
our efforts on the \textit{rpsB-gfp} strain (see \ref{S6_Text} for more details).

@@

All experiments were carried out in sterilized media. Sterilization
was performed by autoclaving for instruments and filtration for solutions
(0.2 $\mu$m). During cloning, bacteria were grown in 20-mL Erlenmeyer
flasks filled with LB or spread on Petri dishes with LA. In all the
experiments, we used M9 minimal medium supplemented with trace elements,
thiamine and a carbon source. The full recipe is reproduced below.
Numbers in squared brackets are characteristics of the stock solution.
All stock solutions were stored at room temperature, except for FeSO$_{4}$~{[}30
g.L\textsuperscript{-1}{]} and Thiamine~{[}10 g.L\textsuperscript{-1}{]},
which were stored respectively at -20$^{\circ}$C and 4$^{\circ}$C.

\begin{center}
\begin{tabular}{|l r|}
  \hline
  \multicolumn{2}{|c|}{\textbf{M9 medium (100 mL)}}\\
  \hline
  CaCl2 [1 mol.L\textsuperscript{-1}] & 10 $\mu$l\\
  MgSO$_4$ [1 mol.L\textsuperscript{-1}] & 200 $\mu$l\\
   5x Salts & 20 mL \\
  Traces elements & 90 $\mu$l \\
  FeSO$_4$ [30 g.L\textsuperscript{-1}] & 10 $\mu$l \\
  Thiamine [10 g.L\textsuperscript{-1}] & 50 $\mu$l \\
  Carbon source & at will \\
  H$_2$0 [18.2 M$\Omega$.cm] & to 100 mL \\
  \hline
\end{tabular}
\begin{tabular}{|l r|}
  \hline
  \multicolumn{2}{|c|}{\textbf{Traces elements (0.9 mL)}}\\
  \hline
  H$_2$0 [18.2 M$\Omega$.cm] & 200$\mu$l\\
  Na$_2$EDTA 2H$_2$O [150 g.L\textsuperscript{-1}] & 100$\mu$l\\
  ZnSO$_4$ 7H$_2$O [45 g.L\textsuperscript{-1}] & 100$\mu$l\\
  CoCl$_2$ 6H$_2$O [3 g.L\textsuperscript{-1}] & 100$\mu$l\\
  MnCl$_2$ 4H$_2$O [10 g.L\textsuperscript{-1}] & 100$\mu$l\\
  H$_3$BO$_3$ [10 g.L\textsuperscript{-1}] & 100$\mu$l\\
  Na$_2$MoO$_4$ 2H$_2$O [4 g.L\textsuperscript{-1}] & 100$\mu$l\\
  CuSO$_4$ 5H$_2$O [3 g.L\textsuperscript{-1}] & 100$\mu$l\\
  \hline
\end{tabular}
\begin{tabular}{|l r|}
    \hline
  \multicolumn{2}{|c|}{\textbf{5x Salts (100 mL)}}\\
  \hline
  Na$_2$HPO$_4$2H$_2$0 & 4.25 g \\
  KH$_2$HPO$_4$ & 1.5 mg \\
  NaCl & 0.25 g\\
  NH$_4$Cl& 0.5 g\\
  H$_2$0 [18.2 M$\Omega$.cm] & to 100 mL \\
  \hline
\end{tabular}
\end{center}

A large quantity of M9 was prepared several days before the experiment,
and stored at 4$^{\circ}$C. It contained all the necessary components
except FeSO$_{4}$, thiamine, and carbon sources, which were added
just before inoculation of the pre-culture. Except where stated otherwise,
the growth temperature was 37$^{\circ}$C. M9 Acetate contains 0.2\%
acetate (in mass of C$_{2}$H$_{3}$O$_{2}$ per mass of solution),
and M9 Glucose contains 0.2\% glucose (in mass of D-(+)-Glucose per
mass of solution).

Four days before the measurements, the glycerol stock containing the
\textit{rpsB-gfp} strain was spread on a Petri dish of LA, and incubated
at 37$^{\circ}$C. On the following day, an isolated colony was
used to inoculate a cotton-plugged, sterile flask containing 20 mL
of M9 acetate (Fig.~\ref{fig:experiment_schema}). The time of inoculation
was calculated such that the culture obtains an OD of 0.3-0.4 OD at
the beginning of the experiment. This optical density corresponds
to a culture in mid-exponential growth phase.

At the day of the experiment, the pre-culture was concentrated by centrifugation
and re-suspended in 5 mL of M9 acetate, supplemented with 50~mg.mL\textsuperscript{-1}
BSA (passivation buffer) for injection into the microfluidic device.
Channels were filled by diffusion until most of them contained cells
($\sim$1 hour). Data acquisition started after a constant medium
flow was successfully obtained ($\sim$1 hour, depending on the quality
of the microfluidics device).

@@

The device was connected using 0.023\textquotedbl{} inner diameter
polyethylene tubes to a waste and a sterile bottle containing 200
mL of growth medium, which was enough for several days of acquisition.
A microfluidic pump (Elvesys) containing an output flow sensor module
was plugged to the medium bottle and a pressure up to 2~bars was
applied to the bottle. The output flow was set to 50 $\mu$L.min$^{-1}$.

@@

In order to correct for the drift of the device, we calculated
the offset between consecutive images of a given field using cross-correlation
in the Fourier space~\cite{guizar-sicairos_efficient_2008} (see
Scikit-Image documentation at \cite{skimage_cross-correlation}).
All images for a field were thereby aligned to the first acquisition
by a simple translation.

@@

The autofluorescence of the bacteria and the background fluorescence
of the medium were supposed to be negligible and were thus not corrected.
As a control, the same device was used to image bacteria in stationary
phase that do not produce any fluorescent protein. They were indistinguishable
from the camera noise. Of course, using this control we can not affirm
that the autofluorescence is also negligible when the bacteria are
actively growing. Furthermore, the autofluorescence could change in
different media of interest, and may even be different in steady state
and during growth transitions. An independent estimate of the autofluorescence
is therefore difficult to obtain. In the section~\ref{sec:chap3_discussion},
we discuss possible improvements of the experimental setup that would
allow to dynamically co-estimate the autofluorescence with the ribosome
abundance in a single experiment.

@@

The data of the length and the RFU/pixel of each bacteria were analyzed
using Python~3.5.2. In particular, we manipulated the data using
Pandas~0.19.1, Pykalman~0.9.5 and the Curves submodule of Wellfare~0.1.1.

Historical details about the Kalman smoothing procedure are reported
in section~\ref{sec:res_kalman} of the main text. We used the Additive
Unscented Kalman Filter implementation of the Pykalman python module.
This class is reported to be more stable and computationally efficient
with non-linear problems containing additive noise.

As described in section~\ref{sec:res_kalman}, the reconstruction
of the growth rate was done on continuous portions of the length,
\textit{i.e.} between cell divisions. The full problem, as described
in Eq.~\ref{eq:full_mu_prob}, is reproduced below for clarity: 
\begin{eqnarray*}
\dot{V_{\lambda}}(t) & = & \mu(t)\cdot V_{\lambda}(t),\\
\dot{\mu}(t) & = & v(t),\\
\dot{v}(t) & = & w(t),
\end{eqnarray*}
with the measurement model: 
\[
L(t_{k})=V_{\lambda}(t_{k})+\epsilon_{k},
\]
and the initial conditions
\[
V_\lambda (0) = V_{\lambda,0}, \;\;\;\; \mu (0) = \mu_0 , \;\;\;\; v(0) = v_0.
\]
The parameters used as priors for the reconstruction of $\mu$ are
described below. We used an observation variance of 9 pixels\textsuperscript{2}
for the length $L$. The transition variance $\theta^{2}$ (\textit{a.k.a.}
the smoothing factor for $\mu$) is fixed at $10^{-8}$~min\textsuperscript{-6}
for the entire time series. Inheritance between mother and daughter
cells is taken into account by systematically choosing an initial
mean for $\mu_0$ equal to the last estimated value for $\mu$, and for $V_{\gamma,0}$ to half the
last estimated value for $V_{\gamma}$ (as expected for a completely
symmetrical division). At the beginning of the experiment, when no
mother cells are available, these values were fixed to 15~pixels
for $V_{\gamma}$ and 0.004~min\textsuperscript{-1} for $\mu$.
The variances associated with these means are 16~pixels\textsuperscript{2}
and $10^{-4}$~min\textsuperscript{-2}, respectively, for $V_{\gamma,0}$
and $\mu_0$. The initial mean of $v$ is always taken as null, with
an initial variance of $10^{-8}$~min\textsuperscript{-6}. All the
cross-covariances are set to 0 because the system variables are independent
by construction.

For the reconstruction of gene activity, we had to cope with the large
gap in data acquisition between -720 and -150 min. Reconstruction
was then done independently on the continuous sections before and
after this gap. The full problem for the estimation of $u=\gamma\alpha/\beta$,
as described in Eq.~\ref{eq:full_u_prob}, is reproduced here for
clarity: 
\begin{eqnarray*}
\dot{r_{\gamma}}(t) & = & \hat{\mu}(t)\cdot u(t)-\hat{\mu}(t)\cdot r_{\gamma}(t),\\
\dot{u}(t) & = & v(t),\\
\dot{v}(t) & = & w(t),
\end{eqnarray*}
with the measurement model 
\[
F(t_{k})=r_{\gamma}(t_{k})+\eta_{k},
\]
and the initial conditions
\[
r_\gamma (0) = r_{\gamma,0}, \;\;\;\; u (0) = u_0 , \;\;\;\; v(0) = v_0,
\]
where $\hat{\mu}$ is the estimate of $\mu$ above. In both case,
we used as prior an observation variance of 800.4 RFU\textsuperscript{2}
for $F$. The transition variance $\theta^{2}$ (\textit{a.k.a.} the
smoothing factor for $\gamma\alpha/\beta$) is fixed at $10^{2}$~RFU\textsuperscript{2}.min\textsuperscript{-4}.
The initial state means used are 600~RFU for $r_{\gamma,0}$ and 1000~RFU
for $u_0$. The variances associated with these means
are purposely large and fixed at 10\textsuperscript{6}~RFU\textsuperscript{2}
for $r_{\gamma,0}$ and $u_0$. The initial mean for $v_0$
is always taken as null, with an initial variance of $10^{-8}$~RFU.min\textsuperscript{-2},
imposing a null second derivative for the reconstructed signal $\gamma\alpha/\beta$.
Here again, all the cross-covariances are set to 0.

All the parameters cited in this section were chosen through trial
and error, and are therefore certainly optimizable. Possible improvements
are discussed in Section~\ref{sec:chap3_discussion}.
